---
title: "Penguin Flipper Lengths ANOVA"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

```

1. Load packages and read in the penguins.csv data

Packages: tidyverse, car, onewaytests
```{r, include=FALSE}
library(tidyverse)
library(car)
library(onewaytests)

penguins <- read_csv("penguins.csv")
```

Our question: Is there a significant difference in mean flipper lenghts for female Chinstrap, Gentoo, and Adelie Penguins near Palmer Archipelago?

Maybe I can use a one-way ANOVA...
- Independent observations
- Equal variance (Levene's test)
- Normality (histograms/qqplots)

```{r}

counts <- with(penguins, table(Species, Sex))

penguins2 <- penguins %>%
  mutate(
    common_name =
      case_when(
        Species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap",
        Species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo",
        Species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie"
      )
  ) %>% 
  select(-Species) %>% 
  filter(Sex == "FEMALE")
```

Exploratory graphs and Levene's test for equal variances

```{r}

hists <- ggplot(penguins2, aes(x = `Flipper Length (mm)`)) +
  geom_histogram(aes(fill = common_name, scale = "free")) +
  facet_wrap(~ common_name, scale = "free")

hists

qqplot <- ggplot(penguins2, aes(sample = `Flipper Length (mm)`)) +
  geom_qq(aes(color = common_name)) +
  facet_wrap(~common_name, scale = "free")

qqplot

```


Evaluate variances and perform a Levene's Test for equal variances

```{r}
# if the largest sample variances is < 4x greater than the smallest smaple variance, then these are "close enough" to consider variacnes equal

variances <- penguins2 %>% 
  group_by(common_name) %>% 
  summarize(
    variance = var(`Flipper Length (mm)`)
  )

variances # variance units will be in mm^2... the largeset variance (33.1) is definitely not more than 4x greater than the smallest variance (15.2)

#Levene's Test
# leveneTest function in the "car" packages

levene_penguin <- leveneTest(`Flipper Length (mm)` ~ common_name, data = penguins2)

levene_penguin

# H0: No difference in variances (variances are equal)
# HA: Variances are NOT equal

# There is no significant difference in variances across species
```

One-way ANOVA for penguin flipper lengths by species (common_name)

```{r}

# H0: The means across all groups are equal
# HA: At least two group mean differ significantly

penguins_aov <- aov(`Flipper Length (mm)` ~ common_name, data = penguins2)
summary(penguins_aov)

# Post-hoc testing by Tukey's HSD

# H0: No significant difference in pairwise means
# HA: There IS a significant difference in pairwise means

penguin_ph <- TukeyHSD(penguins_aov)
penguin_ph

# All p-values for each species is very low, meaning there are differences among all of them
```

Mean flipper lenghts of female penguins differed significantly for all species studied (one-way anova, F(df) = statistic, p <0.001, $\alpha$ = 0.05 with post-hoc Tukey's HSD $\alpha$ = 0.05).